---
title: "Lab 01 - Hello R"
author: "Cynthia Jiao"
date: "01/19/2025"
output: 
  html_document: 
    fig_height: 4
    fig_width: 9
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```

## Exercises

### Exercise 1

This data frame has 1846 rows and 3 columns (variables).

```{r}
?datasaurus_dozen

datasaurus_dozen %>%
  count(dataset) %>%
  print(13)
```

### Exercise 2

Plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```
Just looking at the correlation coefficient, there seems to be a very weak negative correlation (-0.06). However, the plot does NOT show a linear relationship between x and y. For example, within the range of x = 20 to 40, when x increases, y increases. But within the range of x = 70 to 100, when x increases, y decreases. A correlation coefficient does not necessarily mean that there is a linear relationship. Always plot the data first before running a linear correlation.


### Exercise 3

Plot the star dataset:

```{r plot-star, fig.height=6, fig.width=6}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()

```

Calculate the x and y correlation:

```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```
The correlatin coefficient is -0.063, which is slightly weaker than the coefficient of dino dataset. The star plot does NOT show a linear relationship between x and y. For example, when x ranges from 50 to 60, there is a positive relationship between x and y, but when x ranges from 60 to 70, the relationship turns negative. 


### Exercise 4

Plot the circle dataset:

```{r plot-circle, fig.height=6, fig.width=6}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()

```

Calculate the x and y correlation:

```{r cor-circle}
circle_data %>%
  summarize(r = cor(x, y))

```
The correlatin coefficient is -0.068, which is slightly stronger than the coefficient of dino dataset. The circle plot does NOT show a linear relationship between x and y. For example, when x ranges from 20 to 50, as x increases, y can have multiple values and both increase or decrease, which does not represent a consistent linear relationship.


### Exercise 5

Combining all graphs together:

```{r plot-all, fig.height=8, fig.width=6}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~dataset, ncol = 3) +
  theme(legend.position = "none")

datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(13)
```


